Hydrodynamics and rheology of a vesicle doublet suspension

### 2019-07-18 vesicle_code/docs/adhesion/adhesion.tex
first draft of reviewer 2 done

In a quiescent environment, sub-micron size vesicles are found to form a
doublet due to their van der Waals' attractive
interactions~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}.
Due to the strong van der Waals' adhesive force, the vesicles are far
from spherical shape and the membrane is almost flat in the ``contact"
region. Gires {\em et al.}~used small-deformation analysis to
investigate the hydrodynamic interactions between two vesicles in a
planar shear flow with a long separation
distance~\cite{GiresDankerMisbah2012_PRE}.  \todo[inline]{Also cite
Gires et al.~Physics of Fluids 26, 013304, 2014} They found that the
vesicle interaction could be either attractive or repulsive depending on
the organization of the two vesicles relative to the shear flow. To the
best of authors' knowledge, the effects of close-range vesicle adhesion
on their hydrodynamics under an external flow have not been studied and
quantified. The goal of this work is to use state-of-the-art boundary
integral simulations to numerically investigate the dynamics of two
vesicles in both planar shear flow and extensional flow for a wide range
of vesicle shapes and adhesive strength and distance.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in a planar extensional flow} 
\label{sec:eflow} 
\todo[inline]{Do we discuss the slow tank treading motion when the
vesicles reach steady state.  Reviewer 2 had a comment about this}

@@

\begin{figure}[htp]
  \begin{tabular}{@{}p{0.3\linewidth}@{\quad}p{0.3\linewidth}@{\quad}p{0.3\linewidth}@{}}
  \subfigimg[width=\linewidth]{(a)}{figs/adR4em1adS7em1_ra070_inclinationAngle.pdf} &
  \subfigimg[width=\linewidth]{(b)}{figs/adR4em1adS7em1_finalInclinationAngle.pdf} &
  \subfigimg[width=\linewidth]{(c)}{figs/extensional_adR4em1adS7em1_phaseDiagram.pdf}
  \end{tabular}
%  \includegraphics[height=0.26\textwidth]{figs/adR4em1adS7em1_ra070_inclinationAngle.pdf}
%  \includegraphics[height=0.26\textwidth]{figs/adR4em1adS7em1_finalInclinationAngle.pdf}
%  \includegraphics[height=0.26\textwidth]{figs/extensional_adR4em1adS7em1_phaseDiagram.pdf}
  \caption{\label{fig:extensionalInclinationAngle} (a) The inclination
  angle of a doublet formed by vesicles of reduced area $\Delta A = 0.7$
  in an extensional flow with an extension rate $\dot\epsilon$.  For all
  extension rates, the vesicle's orientation angle tends to a value in
  $(\frac{\pi}{2},\pi)$. (b) The final inclination angle of the doublets
  for a variety of reduced areas.  At larger extension rates, the
    doublet is broken by the flow.  (c) A phase diagram of the long-time
    behavior of a fluid trap formed by two adhering vesicles in an
    extensional flow.  Reduced areas and extension rates that do not
    form a doublet are marked in red, and those that do form a doublet
    are marked in blue.  For the blue marks, the size is proportional to
    the final inclination angle.  Squares indicate that the doublet had
    not tilted to an equilibrium inclination angle in the given time
    horizon.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in a shear flow}
\label{sec:sflow} 
We consider two vesicles suspended in the shear flow $\uu =
\dot\gamma(y,0)$, where $\dot\gamma$ is the shear rate.  To be
consistent with simlulations in previous sections, the vesicles have a
length of $2\pi$ for all reduced areas.  The initial placement of the
vesicles is chosen so that the flow dives them towards one another.  In
the absence of adhesion, once the vesicles are sufficiently close, they
are deflected to opposite sides of the $y$ axis, pass one another, and
then separate.  However, in the presence of adhesion, the vesicles can
form a doublet for certain values of the separation distance $\delta$,
Hamaker constant $\mathcal{H}$, shear rate $\dot\gamma$, and reduced
area $\Delta A$.  If a doublet does form, then it undergoes a periodic
motion with a period that depends on the same parameters.

@@

Next, for several shear rates, we determine the critical Hamaker
constant, $\mathcal{H}_c$, that determines if the vesicles form a
doublet or separate.  Figure~\ref{fig:sflow_phase_diagram} indicates
whether a doublet with $\Delta A = 0.9$ is formed (blue) or not (red)
for two different separation distances.  The plot includes a line of
best fit between the stable and unstable regions.  Since the transition
is linear for both separation distances, this indicates that a
dimensionless number involving the ratio of the Hamaker constant and
shear rate could be used to determine if a doublet forms or not.

>>>>

In a quiescent environment, sub-micron size vesicles are found to form a
doublet due to their van der Waals' attractive
interactions~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}.
Due to the strong van der Waals' adhesive force, the vesicles are far
from spherical shape and the membrane is almost flat in the ``contact"
region. Gires {\em et al.}~used small-deformation analysis to
investigate the hydrodynamic interactions between two vesicles in a
planar shear flow with a long separation
distance~\cite{GiresDankerMisbah2012_PRE}.  They found that the vesicle
interaction could be either attractive or repulsive depending on the
organization of the two vesicles relative to the shear
flow~\cite{gir-sri-mis-pod-cou2014}. To the best of authors' knowledge,
the effects of close-range vesicle adhesion on their hydrodynamics under
an external flow have not been studied and quantified. The goal of this
work is to use state-of-the-art boundary integral simulations to
numerically investigate the dynamics of two vesicles in both planar
shear flow and extensional flow for a wide range of vesicle shapes and
adhesive strength and distance.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in a planar extensional flow} 
\label{sec:eflow}

@@

\begin{figure}[htp]
  \begin{tabular}{@{}p{0.3\linewidth}@{\quad}p{0.3\linewidth}@{\quad}p{0.3\linewidth}@{}}
  \subfigimg[width=\linewidth]{(a)}{figs/adR4em1adS7em1_ra070_inclinationAngle.pdf} &
  \subfigimg[width=\linewidth]{(b)}{figs/adR4em1adS7em1_finalInclinationAngle.pdf} &
  \subfigimg[width=\linewidth]{(c)}{figs/extensional_adR4em1adS7em1_phaseDiagram.pdf}
  \end{tabular}
%  \includegraphics[height=0.26\textwidth]{figs/adR4em1adS7em1_ra070_inclinationAngle.pdf}
%  \includegraphics[height=0.26\textwidth]{figs/adR4em1adS7em1_finalInclinationAngle.pdf}
%  \includegraphics[height=0.26\textwidth]{figs/extensional_adR4em1adS7em1_phaseDiagram.pdf}
  \caption{\label{fig:extensionalInclinationAngle} The long time
  behavior of two vesicles in a fluid trap with Hamaker constant
  $\mathcal{H} = 0.7$ and separation distance $\delta = 0.4$. (a) The
  inclination angle of a doublet formed by vesicles of reduced area
  $\Delta A = 0.7$ in an extensional flow with an extension rate
  $\dot\epsilon$.  For all extension rates, the vesicle's orientation
  angle tends to a value in $(\frac{\pi}{2},\pi)$. (b) The final
  inclination angle of the doublets for a variety of reduced areas.  At
  larger extension rates, the doublet is broken by the flow.  (c) A
  phase diagram of the long-time behavior of a fluid trap formed by two
  adhering vesicles in an extensional flow.  Reduced areas and extension
  rates that do not form a doublet are marked in red, and those that do
  form a doublet are marked in blue.  For the blue marks, the size is
  proportional to the final inclination angle.  Squares indicate that
  the doublet had not tilted to an equilibrium inclination angle in the
  given time horizon.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in a planar shear flow}
\label{sec:sflow} 
We consider two vesicles suspended in the planar shear flow $\uu =
\dot\gamma(y,0)$, where $\dot\gamma$ is the shear rate.  To be
consistent with simulations in previous sections, the vesicles have a
length of $2\pi$ for all reduced areas.  The initial placement of the
vesicles is chosen so that the flow dives them towards one another in a
head-on collision.  In the absence of adhesion, once the vesicles are
sufficiently close, they are deflected to opposite sides of the $y$
axis, pass one another, and then separate.  However, in the presence of
adhesion, the vesicles can form a doublet for certain values of the
separation distance $\delta$, Hamaker constant $\mathcal{H}$, shear rate
$\dot\gamma$, and reduced area $\Delta A$.  If a doublet does form, then
it undergoes a periodic motion with a period that depends on the same
parameters.

@@

Next, for several shear rates, we determine the critical Hamaker
constant, $\mathcal{H}_c$, that determines if the vesicles form a
doublet or separate.  Figure~\ref{fig:sflow_phase_diagram} indicates
whether a doublet with $\Delta A = 0.9$ is formed (blue) or not (red)
for two different separation distances.  The plot includes a line of
best fit between the stable and unstable regions.  Since the transition
is linear for both separation distances, this indicates that a
dimensionless number involving the ratio of the Hamaker constant and
shear rate could be used to determine if a doublet forms or not.  We
note that the critical Hamaker constant also depends on the reduced
area.  For smaller reduced areas, we expect a smaller contact region to
form and this will lower the Hamaker constant, and this behavior will be
investigated in future work.

### 2019-07-18 vesicle_code/docs/adhesion/revisions/reviewer2.tex
first draft of reviewer 2 done

\noindent
\comment{Concerning the stability and dynamics of clusters of vesicles with
adhesive interaction, there have been at least two papers on clusters
in micro-channel flow (computational and experimental) that should be
cited: Brust et al., Scientific Reports 4, 4348 (2014) and Claveria et
al., Soft Matter. 2016 vol 12(39):8235-8245. About rheology, at least
the historical paper by Chien on the shear-thinning of blood related
to RBC aggregation should be cited : Chien, S., Usami, S., Dellenback,
R.~J., Gregersen, M.~I., Nanninga, L.~B.~\& Guest, M.~M. Blood
viscosity: Influence of erythrocyte aggregation. Science 157, 829–831
(1967).}
\begin{itemize}
  \item The reference of Brust et al.~has been added.
  \item The reference of Claveria et al~Has been added.
  \item The reference of Chien et al.~has been added.
\end{itemize}

\noindent
\comment{In addition to ref [54], Gires et al. also published experimental
measurements of hydrodynamic interactions between vesicles in Gires et
al., Phys.~Fluids 26, 013304 (2014)}
\begin{itemize}
  \item rebuttal
\end{itemize}

@@

\noindent
\comment{ Figs 9(right) and 12(right): the stability diagrams are presented
in different parameter spaces for the extensional and shear flow
cases. It would have been interesting to see e.g. the diagram in the
interaction energy vs. flow strength $\chi$ for both cases, for
comparison.}
\begin{itemize}
  \item rebuttal
\end{itemize}

>>>>

\noindent
\comment{Concerning the stability and dynamics of clusters of vesicles with
adhesive interaction, there have been at least two papers on clusters
in micro-channel flow (computational and experimental) that should be
cited: Brust et al., Scientific Reports 4, 4348 (2014) and Claveria et
al., Soft Matter. 2016 vol 12(39):8235-8245. About rheology, at least
the historical paper by Chien on the shear-thinning of blood related
to RBC aggregation should be cited : Chien, S., Usami, S., Dellenback,
R.~J., Gregersen, M.~I., Nanninga, L.~B.~\& Guest, M.~M. Blood
viscosity: Influence of erythrocyte aggregation. Science 157, 829–831
(1967).}
\begin{itemize}
  \item The reference of Brust et al.~has been added.
  \item The reference of Claveria et al~has been added.
  \item The reference of Chien et al.~has been added.
\end{itemize}

\noindent
\comment{In addition to ref [54], Gires et al. also published experimental
measurements of hydrodynamic interactions between vesicles in Gires et
al., Phys.~Fluids 26, 013304 (2014)}
\begin{itemize}
  \item The reference of Gires et al.~has been added
\end{itemize}

@@

\noindent
\comment{ Figs 9(right) and 12(right): the stability diagrams are presented
in different parameter spaces for the extensional and shear flow
cases. It would have been interesting to see e.g. the diagram in the
interaction energy vs. flow strength $\chi$ for both cases, for
comparison.}
\begin{itemize}
  \item For the extensional example, the reduced area has a large effect
    on determining if the doublet is broken or not, and on the
    inclination angle if the doublet is not broken.  Therefore, we have
    elected to investigate the effect of the reduced area for this
    example.

  \item For the shear example, a smaller reduced area will increase the
    size of the contact area, and this will determine the critical
    Hamaker constant required to form a doublet.  However, to avoid
    diverting the focus of the paper, we have elected to divert an
    investigation of this behavior to future work.
\end{itemize}
